### 2016-12-06 Paper_v9.tex
fig order fix

JuPOETs generated a population of probable signaling models which captured the multiple phases of EMT induction (Fig. \ref{fg:F2}).
JuPOETs sampled well over 10$^4$ probable models during each stage of the cross-validation using global random sampling.
From this analysis, N $\simeq$ 1400 models were selected for further analysis.
The selected models all had the same possible molecular connectivity, but different values for model parameters.
Transcription and translation rates, as well as mRNA and protein degradation terms, were set using physical values from the literature \cite{Milo:2010aa},
and allowed to vary by a scaling factor, see methods.
Model selection was based upon Pareto rank, the prediction and training error across all objectives.
The model population recapitulated key signaling events following TGF-$\beta$ exposure.
We subdivided the response to TGF-$\beta$ exposure into two phases.
First, TGF-$\beta$1/2 signaling initiated a program which downregulated E-cadeherin expression in a MAPK dependent manner while
simultaneously upregulating TGF-$\beta$3 expression. Second, TGF$-\beta$3 secretion initiated an autocrine feedback which upregulated the expression of
mesenchymal markers such as Vimentin and key upstream transcription factors such as LEF-1 in a SMAD dependent manner.
TGF-$\beta$3 expression was also able to sustain $\beta$-catenin release by inhibiting its sequestration by the APC complex through PI3K mediated GSK3,
which was captured by the model (Fig. \ref{fg:F4}B).
Each phase involved the hierarchal expression and/or post-translational modification of several key transcription factors.
During the first phase, stimulation with TGF-$\beta$1/2 (10 a.u.) activated both the SMAD and MAPK pathways.
MAPK activation resulted in the phosphorylation of the transcription factor activator protein 1 (AP-1), which in-turn upregulated the expression of Snail, a well established transcriptional repressor (Fig. \ref{fg:F2}A).
Snail expression was MAPK-dependent; the MEK inhibitor U0126 blocked AP-1 activation and Snail expression following TGF-$\beta$1/2 exposure (Fig. \ref{fg:F2}A, Lane 3).
Overexpression of either Snail or Slug upregulated TGF-$\beta$3 expression (Fig. \ref{fg:F2}C) while simultaneously downregulating E-cadeherin expression (Fig. \ref{fg:F2}F).
During the second phase, TGF$-\beta$3 secretion and the subsequent autocrine signaling resulted in the upregulation of mesenchymal marker expression.
The TGF$-\beta$3 induced gene expression program involves a complex hierarchy of transcriptional and post-translational regulatory events.
Absence of E-cadherin indirectly promoted TGF$-\beta$3 expression through the $\beta$-catenin/TCF4 complex following Snail or Slug expression (Fig. \ref{fg:F2}C, Lane 2 or 3).
Conversely, over-expression of E-cadherin inhibited the TGF$-\beta$3 autocrine production by sequestering cytosolic $\beta$-catenin, thereby blocking EMT (Fig. \ref{fg:F2}C, Lane 4 or 5).
TGF$-\beta$3 signaled through the Smad pathway to regulate LEF-1 expression and downstream target EMT genes (Fig. \ref{fg:F2}G).
TGF$-\beta$3 (10 a.u.) in combination with downstream inhibitors (DN-Smad4 and DN-LEF-1) completely inhibited Vimentin expression, while elevating E-cadherin expression (Fig. \ref{fg:F2}H,I).

@@

\subsection*{Identification of a novel LEF-1 regulator}
During model identification, we found that consistent TGF-$\beta$ induced EMT from a stable epithelial cell population required an additional regulatory protein.
This protein, which we called hypothetical regulator 1 (YREG1), was required to mediate between SNAIL/SLUG transcriptional activity
and the upregulation of LEF-1 expression following TGF-$\beta$1/2 exposure.
SNAIL/SLUG are well known transcriptional repressors \cite{Hemavathy:2000aa,Hemavathy:2000ab,Dhasarathy:2011aa}, although there are a few studies
which suggest that at least SNAIL can also act as a transcriptional activator \cite{Guaita:2002aa}.
In the model, we assumed the expression of SNAIL/SLUG was likely regulated by AP1/SP1 \cite{Jackstadt:2013aa}.
Thus, upon receiving direct SNAIL/SLUG and TGF-$\beta$3 signals, the model predicted enhanced SNAIL/SLUG expression, consistent with experimental observations.
TGF-$\beta$1/2 stimulation also induces LEF-1 expression. However, literature evidence suggested that LEF-1 expression was not strongly dependent upon AP1/SP1 activity \cite{Eastman:1999aa}.
Thus, either SNAIL/SLUG are acting as inducers (contrary to substantial biochemical evidence) or, they are repressing the expression of an intermediate repressor.
Given the biochemical evidence supporting SNAIL/SLUG as repressors, we created the hypothetical YREG1 repressor whose
expression is downregulated by SNAIL/SLUG. The literature data therefore suggested that YREG1 had two transcriptional targets, LEF-1 and TGF-$\beta$3.
By adding this regulator, our simulations became consistent with training and literature data.
Medici et al. suggested that feedback between $\beta$-catenin and LEF-1 was likely, although this feedback had yet to be identified \cite{Medici:2008fk}.
Low levels of YREG1 expression were present in all simulations to regulate the formation of the $\beta$-catenin-LEF-1 complex.
To test the effect of YREG1 on the epithelial population, we conducted over-expression and knockdown simulations on untreated cells (Fig. \ref{fg:F4}C and \ref{fg:F4}D).
In the absence of YREG1, the population of models failed to consistently to retain a stable epithelial state (Fig. \ref{fg:F4}D).
Conversely, YREG1 amplification revealed an enhanced epithelial phenotype, while some inherently transformed cells moved towards a hybrid phenotype  (Fig. \ref{fg:F4}C).
Elevated YREG1 repressed LEF-1 and TGF-$\beta$3 expression, thereby not allowing free $\beta$-catenin to form the $\beta$-catenin-LEF-1 complex, or TGF-$\beta$3 induced SMAD activation.
Taken together, low YREG1 expression was required for the maintenance of a stable epithelial phenotype that was simultaneously inducible across TGF-$\beta$1/2, TGF-$\beta$3 and SNAIl/SLUG transfection.

@@

\begin{figure}
%\includegraphics [width=1.0\linewidth] {./figs/Fig4_mechanism.pdf}
\caption{Analysis of underlying signaling responses.
(A) We examined the distribution of NFATc1 and AP1/SP1 in cells containing the hybrid phenotype (VEGF-A + TGF$-\beta$2 case), showing the potential for cells to express both SP1 and NFATc1 in a non exclusive manner.
(B) We fit an additional objective demonstrating the activation of GSK3 through PI3K. The model captured this activation through TGF$-\beta$3 signaling. LY294002 is a PI3K inhibitor.
(C) We identified a novel regulator of LEF1 called YREG1 that allows Snail/Slug to emulate an inducer by repressing YREG1, which was required to stabilize the untreated population.
YREG1 overexpression revealed an enhanced epithelial phenotype, while some inherently transformed cells moved towards a hybrid phenotype.
(D) In the absence of YREG1, most of the population failed to consistently to retain a stable epithelial state.}\label{fg:F4}
\end{figure}

@@

>>>>

JuPOETs generated a population of probable signaling models which captured the multiple phases of EMT induction (Fig. \ref{fg:F2}).
JuPOETs sampled well over 10$^4$ probable models during each stage of the cross-validation using global random sampling.
From this analysis, N $\simeq$ 1400 models were selected for further analysis.
The selected models all had the same possible molecular connectivity, but different values for model parameters.
Transcription and translation rates, as well as mRNA and protein degradation terms, were set using physical values from the literature \cite{Milo:2010aa},
and allowed to vary by a scaling factor, see methods.
Model selection was based upon Pareto rank, the prediction and training error across all objectives.
The model population recapitulated key signaling events following TGF-$\beta$ exposure.
We subdivided the response to TGF-$\beta$ exposure into two phases.
First, TGF-$\beta$1/2 signaling initiated a program which downregulated E-cadeherin expression in a MAPK dependent manner while
simultaneously upregulating TGF-$\beta$3 expression. Second, TGF$-\beta$3 secretion initiated an autocrine feedback which upregulated the expression of
mesenchymal markers such as Vimentin and key upstream transcription factors such as LEF-1 in a SMAD dependent manner.
TGF-$\beta$3 expression was also able to sustain $\beta$-catenin release by inhibiting its sequestration by the APC complex through PI3K mediated GSK3,
which was captured by the model (Fig. \ref{fg:gsk3_alone}).
Each phase involved the hierarchal expression and/or post-translational modification of several key transcription factors.
During the first phase, stimulation with TGF-$\beta$1/2 (10 a.u.) activated both the SMAD and MAPK pathways.
MAPK activation resulted in the phosphorylation of the transcription factor activator protein 1 (AP-1), which in-turn upregulated the expression of Snail, a well established transcriptional repressor (Fig. \ref{fg:F2}A).
Snail expression was MAPK-dependent; the MEK inhibitor U0126 blocked AP-1 activation and Snail expression following TGF-$\beta$1/2 exposure (Fig. \ref{fg:F2}A, Lane 3).
Overexpression of either Snail or Slug upregulated TGF-$\beta$3 expression (Fig. \ref{fg:F2}C) while simultaneously downregulating E-cadeherin expression (Fig. \ref{fg:F2}F).
During the second phase, TGF$-\beta$3 secretion and the subsequent autocrine signaling resulted in the upregulation of mesenchymal marker expression.
The TGF$-\beta$3 induced gene expression program involves a complex hierarchy of transcriptional and post-translational regulatory events.
Absence of E-cadherin indirectly promoted TGF$-\beta$3 expression through the $\beta$-catenin/TCF4 complex following Snail or Slug expression (Fig. \ref{fg:F2}C, Lane 2 or 3).
Conversely, over-expression of E-cadherin inhibited the TGF$-\beta$3 autocrine production by sequestering cytosolic $\beta$-catenin, thereby blocking EMT (Fig. \ref{fg:F2}C, Lane 4 or 5).
TGF$-\beta$3 signaled through the Smad pathway to regulate LEF-1 expression and downstream target EMT genes (Fig. \ref{fg:F2}G).
TGF$-\beta$3 (10 a.u.) in combination with downstream inhibitors (DN-Smad4 and DN-LEF-1) completely inhibited Vimentin expression, while elevating E-cadherin expression (Fig. \ref{fg:F2}H,I).

@@

\subsection*{Identification of a novel LEF-1 regulator}
During model identification, we found that consistent TGF-$\beta$ induced EMT from a stable epithelial cell population required an additional regulatory protein.
This protein, which we called hypothetical regulator 1 (YREG1), was required to mediate between SNAIL/SLUG transcriptional activity
and the upregulation of LEF-1 expression following TGF-$\beta$1/2 exposure.
SNAIL/SLUG are well known transcriptional repressors \cite{Hemavathy:2000aa,Hemavathy:2000ab,Dhasarathy:2011aa}, although there are a few studies
which suggest that at least SNAIL can also act as a transcriptional activator \cite{Guaita:2002aa}.
In the model, we assumed the expression of SNAIL/SLUG was likely regulated by AP1/SP1 \cite{Jackstadt:2013aa}.
Thus, upon receiving direct SNAIL/SLUG and TGF-$\beta$3 signals, the model predicted enhanced SNAIL/SLUG expression, consistent with experimental observations.
TGF-$\beta$1/2 stimulation also induces LEF-1 expression. However, literature evidence suggested that LEF-1 expression was not strongly dependent upon AP1/SP1 activity \cite{Eastman:1999aa}.
Thus, either SNAIL/SLUG are acting as inducers (contrary to substantial biochemical evidence) or, they are repressing the expression of an intermediate repressor.
Given the biochemical evidence supporting SNAIL/SLUG as repressors, we created the hypothetical YREG1 repressor whose
expression is downregulated by SNAIL/SLUG. The literature data therefore suggested that YREG1 had two transcriptional targets, LEF-1 and TGF-$\beta$3.
By adding this regulator, our simulations became consistent with training and literature data.
Medici et al. suggested that feedback between $\beta$-catenin and LEF-1 was likely, although this feedback had yet to be identified \cite{Medici:2008fk}.
Low levels of YREG1 expression were present in all simulations to regulate the formation of the $\beta$-catenin-LEF-1 complex.
To test the effect of YREG1 on the epithelial population, we conducted over-expression and knockdown simulations on untreated cells (Fig. \ref{fg:F4}B and \ref{fg:F4}C).
In the absence of YREG1, the population of models failed to consistently to retain a stable epithelial state (Fig. \ref{fg:F4}B).
Conversely, YREG1 amplification revealed an enhanced epithelial phenotype, while some inherently transformed cells moved towards a hybrid phenotype  (Fig. \ref{fg:F4}C).
Elevated YREG1 repressed LEF-1 and TGF-$\beta$3 expression, thereby not allowing free $\beta$-catenin to form the $\beta$-catenin-LEF-1 complex, or TGF-$\beta$3 induced SMAD activation.
Taken together, low YREG1 expression was required for the maintenance of a stable epithelial phenotype that was simultaneously inducible across TGF-$\beta$1/2, TGF-$\beta$3 and SNAIl/SLUG transfection.

@@

\begin{figure}
%\includegraphics [width=1.0\linewidth] {./figs/Fig4_mechanism.pdf}
\caption{Analysis of underlying signaling responses.
(A) We examined the distribution of NFATc1 and AP1/SP1 in cells containing the hybrid phenotype (VEGF-A + TGF$-\beta$2 case), showing the potential for cells to express both SP1 and NFATc1 in a non exclusive manner.
(B) In the absence of YREG1, most of the population failed to consistently to retain a stable epithelial state.}\label{fg:F4}
(C) We identified a novel regulator of LEF1 called YREG1 that allows Snail/Slug to emulate an inducer by repressing YREG1, which was required to stabilize the untreated population.
YREG1 overexpression revealed an enhanced epithelial phenotype, while some inherently transformed cells moved towards a hybrid phenotype.
\end{figure}

@@

\begin{figure}
%\includegraphics [width=1.0\linewidth] {./figs/Fig_gsk3_sup.pdf}
\caption{We fit an additional objective demonstrating the activation of GSK3 through PI3K. The model captured this activation through TGF$-\beta$3 signaling. LY294002 is a PI3K inhibitor.} 
\label{fg:gsk3_alone}
\end{figure}
